---
title: "plottyArt_PRJ_two"
author: "Brady M. Chisholm"
date: "2024-09-27"
output: html_document
---

Nonfunction version for testing / updates
```{r}
graphics.off() # close all figs, but generate new ones 
  library(ggplot2)
  # Generate sequenced values based on input
  startSeq = -1000
  endSeq = 1000
  countMe = .05 # small step size looks smoother on plot 
  plottingSeq = seq(startSeq,endSeq,countMe) # ways to manipulate these numbers so they have an interesting pattern / relationship 
  # generate the 'funky' y coordinate from my formula 
  
  formulaVals <- (plottingSeq+(plottingSeq^3))^4  
  # define custom colors 
  customPalette <- c("#EDADC7","#D199B6","#C493B0","#A68BA5")
  palette_fn <-colorRampPalette(customPalette) 
  
  # Generate number of colors based on number of points 
  customColors <- palette_fn(length(plottingSeq))
  dataToPlot <- data.frame(x = plottingSeq, y = formulaVals, color = customColors)

  # Generate fig 
  myFig <- ggplot(dataToPlot, aes(x = y, y = x, color = color)) +
    geom_point()+  # Use the first color for all points
    scale_color_identity()+  # Use custom colors directly 
    ggtitle(paste(startSeq, "to", endSeq, "plot")) + # title the plot with number ranges included 
    guides(color = "none") + 
    theme_bw() + # basic black and white theme 
    theme(
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      axis.text = element_blank(),         
      axis.ticks = element_blank(),        
      axis.title = element_blank()         
    ) # close theme 
  # return plot, needs print function  
  print(myFig)
```
#################
Function creation
#################
Plots before were cool, but looked pretty logarthmic. Would like to generate some plots that curve the data and allow for more abstract / random movement. Not sure if this is a ggplot manipulation or some math I need to do on the data. I think seed() function use is the only thing that generates truly random numbers?  

Going to introduce degrees input so I am hopefully able to manipulate the data with some specified curve. 
```{r}
plotMyCustomArtFunctionV_2 <- function(start,end,by){
  library(ggplot2)
  # Generate sequenced values based on input
  plottingSeq = seq(start,end,by) # ways to manipulate these numbers so they have an interesting pattern / relationship 
  # generate the 'funky' y coordinate from my formula 
  formulaVals <- sqrt((plottingSeq+(plottingSeq^3))^2) # boring formula 
  # define custom colors 
  customPalette <- c("#EDADC7","#D199B6","#C493B0","#A68BA5")
  palette_fn <-colorRampPalette(customPalette) 
  
  # Generate number of colors based on number of points 
  customColors <- palette_fn(length(plottingSeq))
  dataToPlot <- data.frame(x = plottingSeq, y = formulaVals, color = customColors)

  # Generate fig 
  myFig <- ggplot(dataToPlot, aes(x = x, y = y, color = color)) +
    geom_point()+  # Use the first color for all points
    scale_color_identity()+  # Use custom colors directly 
    ggtitle(paste(start, "to", end, "plot!")) + # title the plot with num ranges included 
    guides(color = "none") + 
    theme_bw() + 
    theme(
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      axis.text = element_blank(),         
      axis.ticks = element_blank(),        
      axis.title = element_blank()         
    ) # close theme 
  # return plot, needs print function  
  print(myFig)
  
} # function end 
## Call test cases ##
```



