---
title: "rArtPlayground"
author: "Brady M. Chisholm"
date: "2024-09-27"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
###############
Formula Testing
###############
Added startSeq, endSeq, countMe, and the formula into titles in the plot. It is easy to save images of plots, but not the formulas along as I change them. Just record on plot. Improve reproducibility. 
```{r my special fig one}
library(ggplot2)
library(dplyr)
library(crayon)
library(aRtsy)
#####################
# MAIN GRAPH OPTIONS#
#####################
# Generate sequenced values based on input; manipulate numbers to find interesting pattern / relationship
  startSeq = 1
  endSeq = 20000  
  countMeBy = 2
  pixelSize = 1
# small step size smoother on larger sets of data? 
# main need is to test different formulas  
formula_str <- "(sin(x)*(x^2))"

# Generate x values (sequence of numbers)
plottingSeq <- seq(startSeq, endSeq, countMeBy)

# Evaluate the formula for the x values
formulaVals <- eval(parse(text = formula_str), list(x = plottingSeq))

# decide custom colors. Code currently ramps / distributes a color gradient evenly for the n amount of points that exist 
customPalette <-  c("#EDADC7","#A68BA5")
palette_fn <-colorRampPalette(customPalette) 
customColors <- palette_fn(length(plottingSeq))
######KEEP OPEN OR KILLL###  
#graphics.off()
backgroundAspect <- c("#1C0F13","#6E7E85","#B7CECE","#BBBAC6","#E2E2E2")
palette_fn <- colorRampPalette(backgroundAspect)
backgroundAspectRamped <- palette_fn(length(plottingSeq))
###########################  
  # combine all of our values into a data frame to pass to ggPlot
  mySpedData <- data.frame(x = plottingSeq, y = formulaVals, color = customColors)
  # General fig code below. Includes use of custom color var
  myFig <- ggplot(data = mySpedData) +
    color = (customColors) +
    size = (pixelSize) + 
    scale_color_identity() +  # Use the first color for all points
    ggtitle(paste(startSeq, "to", endSeq, "plot.", "Counted By =",countMeBy)) +  #title with num ranges and step size
    labs(subtitle = bquote(.(formula_str))) +  # add formula in subtitle
    guides(color = "none") + 
    theme_bw() +
    theme(
      panel.background = element_rect(fill = backgroundAspectRamped, colour = "#E2E2E", linewidth = 3),
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      axis.text = element_blank(),         
      axis.ticks = element_blank(),        
      axis.title = element_blank()         
    )
  # display figure
  print(myFig)
```
 
```{r mathart butterfly function}
butterfly_curve <- function(n = 1000, a = 12, b = 12, c = 2, d = 2,
                            e = 4, f = 4, g = 5, h = 5, i = 12, j = 12) {
  result <- data.frame(t = seq(0, a*pi, b*pi/n)) %>%
    dplyr::mutate(x = sin(t) * (exp(cos(t)) - c*cos(e*t) - (sin(t/i))^g),
                  y = cos(t) * (exp(cos(t)) - d*cos(f*t) - (sin(t/j))^h)) %>%
    dplyr::select(-t)
  result
}
```


  if (save ==1){
  saveMyFig <- function(myFunFig){
  #FIXME add saving functionality to save figs with the range and step in title
}}
Formulas to try:


##############################################################
Generate an image of our custom colors on a evenly distributed gradient based on the length of our sequence
```{r palette image}
# generate image of our color on a gradiant 
  imgMatrix <- matrix(1:length(plottingSeq), ncol = 1)
# generate our random image 
image(
  x = imgMatrix, 
  col = customColors, 
  useRaster = TRUE,
  axes = FALSE
)
```

##############################################################
Manipulate dims of the data. Currently identical to the code above. Should make proper edits to make sure no important vars overlap or anything... 
```{r my special fig two, dim change}
library(ggplot2)
library(dplyr)
library(crayon)
#library(doMC) # crash? 
#registerDoMC()
#getDoParWorkers(4) #This lists how many workers you have (hopefully more than 1!)
#####################
# MAIN GRAPH OPTIONS#
#####################
  # manipulate these numbers so they have interesting pattern / relationship
  # Generate sequenced values based on input
  startSeq = -500
  endSeq = 500
  countMe = .5 
  pixelSize = 2 
# small step size looks smoother on plot?? 
# main need is to test different formulas  
formula_str <- "tan(cos(x^2))"  # change formula as needed

#HELPER FUNCTION switch text to mathematical formula. A bit recursive 
generateFormula <- function(formula) {
  function(x) {eval(parse(text = formula))}}
plottingSeq = seq(startSeq,endSeq,countMe) 

# HELPER AI CODE: define formula as a string, create y vals for plot
funkyForm_fn <- generateFormula(formula_str)
formulaVals <- funkyForm_fn(plottingSeq)

# decide custom colors. Code currently ramps / distributes a color gradient evenly for the n amount of points that exist 
  customPalette <- c("#EDADC7","#A68BA5")
  palette_fn <-colorRampPalette(customPalette) 
  customColors <- palette_fn(length(plottingSeq))
  
  #graphics.off()
  
  # combine all of our values into a data frame to pass to ggPlot
  mySpedData <- cbind.data.frame(x = plottingSeq, y = formulaVals, color = customColors)
  # General fig code below. Includes use of custom color var
  myFig <- ggplot(mySpedData) +
    geom_point(aes(x,y, color = color),size = pixelSize)+ 
    scale_color_identity() +  # Use the first color for all points
    ggtitle(paste(startSeq, "to", endSeq, "plot.", "Step =",countMe)) +  #title with num ranges and step size
    labs(subtitle = bquote(.(formula_str))) +  # add formula in subtitle
    guides(color = "none") + 
    theme_bw() +
    theme(
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      axis.text = element_blank(),         
      axis.ticks = element_blank(),        
      axis.title = element_blank()         
    ) 
  print(myFig)
```

Try to plot polar coords. I think regular ggplot & geom_point don't plot these like I want. 
Seems like plots with polar coords may require something stronger like parallel computing. 
```{r plot in polar}
library(ggplot2)
library(dplyr)
library(crayon)
#####################
# MAIN GRAPH OPTIONS#
#####################
  # manipulate these numbers so they have interesting pattern / relationship
  # Generate sequenced values based on input
  startSeq = 1
  endSeq = 1200
  countMe = 1
  pixelSize = .05 
# small step size looks smoother on plot?? 
# main need is to test different formulas  
formula_str <- "(plottingSeq)^.05"  # change formula as needed

#HELPER FUNCTION switch text to mathematical formula. A bit recursive 
generateFormula <- function(formula) {
function(x) {eval(parse(text = formula))}}
plottingSeq = seq(startSeq,endSeq,countMe) 

# HELPER AI CODE: define formula as a string, create y vals for plot
funkyForm_fn <- generateFormula(formula_str)
formulaVals <- funkyForm_fn(plottingSeq)

# custom colors. Code currently ramps / distributes a color gradient evenly for the n amount of points that exist 
  customPalette <- c("#EDADC7","#A68BA5")
  palette_fn <-colorRampPalette(customPalette) 
  customColors <- palette_fn(length(plottingSeq))
  
  #graphics.off()
  
  # wrangle into a data frame to pass to fig
  mySpedData <- cbind.data.frame(x = plottingSeq, y = formulaVals, color = customColors)
  
  # Polar fig code below. Includes use of custom color var
  myFig <- ggplot(mySpedData, aes(x = x, y = y, color = color)) +
  geom_point(size = pixelSize) +  # Add points to plot
  coord_polar() +  # Use polar coordinates
  scale_color_identity() +  # Use custom colors
  ggtitle(paste("Plot", startSeq,'to', endSeq, " ", "Step =", countMe)) +
  labs(subtitle = formula_str) +  # Add formula as subtitle
  guides(color = "none") + 
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )

print(myFig)
```

test out the aRtsy package
source code on GitHub shows ggplot under the hood of this package. Formulas are just much more complex than I thought they would be. 
```{r aRtsy chladni}
library(aRtsy)
# Simple example
#canvas_chladni(colors = colorPalette("origami"))
  #customPalette <- c("#EDADC7","#A68BA5")
  customPalette <- c("#8B1E3F","#3C153B","#89BD9E","#F0C987","#DB4C40")
  palette_fn <- {colorRampPalette(customPalette)}
  customColors <- palette_fn(length(customPalette))
  
set.seed(1)
# Advanced example
canvas_chladni(
  colors = customPalette,
  waves = c(0,3,6,9,12,15,18,21,200),
  warp = 35)
```
 
```{r planets}
canvas_planet(
  colors = customColors,
  threshold = 4,
  iterations = 500, # how many times algorithm runs 
  starprob = 0.001, # probability of a star 
  fade = .05, 
  radius = NULL, # radius of planets. X Y below related  
  center.x = NULL, 
  center.y = -10,
  light.right = TRUE, # light source direction
  resolution = 1500
)
```

This seems like it would be a good function to look at source code for...ggplot back-end 
```{r segments}
canvas_segments(
  colors = customColors, 
  background = 'white', # specify background for the art. null = white? 
  n = 500, # number of line segments to draw 
  p = 0.1, #probability of a vertical line segment 
  H = .1, # scale factor of line segments 
  size = 10) # size of line segments 
```

draw mesh? 
```{r mesh}
library(aRtsy)

canvas_mesh(
  colors = customColors,
  background = "white",
  transform = c("perlin", "fbm", "simplex", "cubic",
                "worley", "knn", "rf", "svm"),
  lines = 50,
  iterations = 500,
  mixprob = .5
)

#ggplot2::ggsave(filename = "mySuperCoolFigure_v1.0.0.jpg", plot = myCoolfig, width = 7, height = 7, dpi = 500)

```
solve color function problem...how to get gradient? 
```{r color gradient trouble shoot}
customPalette <- c("#8B1E3F","#3C153B","#89BD9E","#F0C987","#DB4C40")
  palette_fn <- colorRampPalette(customPalette)
  customColors <- palette_fn(100)

image(1:length(customColors), 1, as.matrix(1:100), col = customGradient, useRaster = TRUE, axes = FALSE)
```

```{r custom color gradient function}
gradientMyCols <- function(img = FALSE, ...){
  # Capture colors as arguments
  colors <- list(...)
  
  # Check if fewer than 2 colors provided
  if(length(colors) < 2){
    stop("You must provide 2 colors")}
  
  if(length(colors) > 8){ # limit max to 8 (currently)
    warning("Only 8 colors will be used")
    colors <- colors[1:8]}

  colors <- unlist(colors) # colors to character vector 
  
  # Create the color gradient
  palette_fn <- colorRampPalette(colors, bias = 1) # grDevices function 
  
  # If img true, generate gradient image
  if (img) {
    # Generate 150 colors for gradient
    gradientColors <- palette_fn(150)
    # Plot the gradient
    image(1:150, 1, as.matrix(1:150), xlab = '', ylab='', col = gradientColors, useRaster = TRUE, axes = FALSE)}
  
    return(palette_fn) }# return color palette function 

myGradient <- gradientMyCols(1,"#8B1E3F", "#3C153B", "#89BD9E")

gradientMyCols(TRUE, "#8B1E3F", "#3C153B", "#89BD9E", "#F0C987", "#DB4C40")
```


```{r GPT ggplot art example}
library(ggplot2)

t <- seq(0, 10*pi, length.out = 1000)
x <- t * cos(t)
y <- t * sin(t)

df <- data.frame(x = x, y = y)

ggplot(df, aes(x = x, y = y)) +
  geom_path(color = "purple", size = 1) +
  theme_void()  # Remove axes and labels

```

```{r GPT grid based art }
library(grid)

grid.newpage()
for (i in seq(0, 1, length.out = 20)) {
  for (j in seq(0, 1, length.out = 20)) {
    grid.circle(x = i, y = j, r = 0.02, gp = gpar(col = rgb(i, j, 0.5), fill = rgb(i, j, 0.5)))
  }
}

```

```{r use math libraries}
# Example: Mandelbrot Set
mandelbrot <- function(c, max_iter = 100) {
  z <- c
  for (i in 1:max_iter) {
    z <- z^2 + c
    if (Mod(z) > 2) return(i)
  }
  return(max_iter)
}

nx <- 800
ny <- 800
x <- seq(-2, 2, length.out = nx)
y <- seq(-2, 2, length.out = ny)
grid <- outer(x, y, FUN = function(x, y) mandelbrot(complex(real = x, imaginary = y)))

image(x, y, grid, col = rainbow(100), useRaster = TRUE, axes = FALSE, xlab = "", ylab = "")

```

```{r turtle in R}
library(TurtleGraphics)

turtle_init(width = 10, height = 10)
turtle_do({
  for (i in 1:36) {
    turtle_forward(50)  # Reduced distance
    turtle_right(170)
  }
})

```

```{r spirographs}
# Example: Hypotrochoid (Spirograph-like patterns)
t <- seq(0, 2*pi, length.out = 1000)

R <- 5   # Radius of the fixed circle
r <- 3   # Radius of the moving circle
d <- 5   # Distance from center of moving circle to the drawing point

x <- (R - r) * cos(t) + d * cos(((R - r) / r) * t)
y <- (R - r) * sin(t) - d * sin(((R - r) / r) * t)

plot(x, y, type = 'l', col = "blue", axes = FALSE, xlab = "", ylab = "", asp = 1)

```

```{r mandelbrot set }
mandelbrot <- function(c, max_iter = 100) {
  z <- 0
  for (i in 1:max_iter) {
    z <- z^2 + c
    if (Mod(z) > 2) return(i)
  }
  return(max_iter)
}

x_vals <- seq(-2, 1, length.out = 500)
y_vals <- seq(-1.5, 1.5, length.out = 500)

mandelbrot_set <- outer(x_vals, y_vals, Vectorize(function(x, y) mandelbrot(complex(real = x, imaginary = y))))

image(x_vals, y_vals, mandelbrot_set, col = terrain.colors(100), axes = FALSE, xlab = "", ylab = "")

```

```{r Lorenz Attractor}
library(deSolve)

#The Lorenz system is a system of differential equations that exhibits chaotic behavior. It’s often visualized as an attractor in 3D space.
lorenz <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dx <- sigma * (y - x)
    dy <- x * (rho - z) - y
    dz <- x * y - beta * z
    list(c(dx, dy, dz))
  })
}

parameters <- c(sigma = 10, rho = 28, beta = 8/3)
state <- c(x = 1, y = 1, z = 1)
times <- seq(0, 100, by = 0.01)

out <- ode(y = state, times = times, func = lorenz, parms = parameters)

# Plotting the Lorenz attractor
plot(out[, "x"], out[, "z"], type = "l", col = "red", xlab = "X", ylab = "Z", main = "Lorenz Attractor")

```

```{r sierpinski triangle}
sierpinski <- function(n, p1, p2, p3) {
  if (n == 0) {
    lines(c(p1[1], p2[1]), c(p1[2], p2[2]))
    lines(c(p2[1], p3[1]), c(p2[2], p3[2]))
    lines(c(p3[1], p1[1]), c(p3[2], p1[2]))
  } else {
    mid1 <- (p1 + p2) / 2
    mid2 <- (p2 + p3) / 2
    mid3 <- (p3 + p1) / 2
    sierpinski(n-1, p1, mid1, mid3)
    sierpinski(n-1, mid1, p2, mid2)
    sierpinski(n-1, mid3, mid2, p3)
  }
}

plot.new()
plot.window(xlim = c(0, 1), ylim = c(0, 1))

sierpinski(5, c(0, 0), c(1, 0), c(0.5, sqrt(3)/2))

```

```{r 3-d plot of mathematical functions}
library(plot3D)

x <- seq(-10, 10, length.out = 100)
y <- seq(-10, 10, length.out = 100)
z <- outer(x, y, function(x, y) sin(sqrt(x^2 + y^2)))

persp3D(x, y, z, col = "blue", theta = 30, phi = 20, ltheta = 120, shade = 0.5, ticktype = "detailed")

```

```{r vorovoi diagrams}
library(deldir)

set.seed(42)
points <- data.frame(x = runif(30), y = runif(30))

vor <- deldir(points$x, points$y)
plot(vor, wlines = "both", lty = 1, lwd = 2, col = "black")
points(points$x, points$y, col = "red", pch = 19)

```

```{r chaos game;triangle}
chaos_game <- function(iter = 10000) {
  vertices <- matrix(c(0, 0, 1, 0, 0.5, sqrt(3)/2), ncol = 2, byrow = TRUE)
  x <- numeric(iter)
  y <- numeric(iter)
  x[1] <- runif(1)
  y[1] <- runif(1)

  for (i in 2:iter) {
    vertex <- vertices[sample(1:3, 1), ]
    x[i] <- (x[i-1] + vertex[1]) / 2
    y[i] <- (y[i-1] + vertex[2]) / 2
  }

  plot(x, y, pch = ".", col = "darkgreen", asp = 1, axes = FALSE)
}

chaos_game()

```

```{r mathematical knots}
library(rgl)

# Trefoil knot parameterization
t <- seq(1,pi, length.out = 100000)

x <- -sin(t-2^3) + 5*sin(2*t)
y <- cos(t+2) - 2*cos(2*t)
z <- -sin(pi*t)
set.seed(1)
# Plot 3D trefoil knot
plot3d(x, y, z, type = "p", radius = 10000, col = "darkgreen", lwd = .05, xlab = "X", ylab = "Y", zlab = "Z", decorate = TRUE)

# Add axes
```

```{r brownian motion}
# Load rgl for 3D plotting GUI 
library(rgl)

# Number of steps
n <- cumsum(1:69)

# Generate Brownian motion (random walk) in 3D
set.seed(3)
x <- cumsum(rnorm(n))
y <- cumsum(rnorm(n))
z <- cumsum(rnorm(n))

# Plot the 3D random walk (Brownian motion)
plot3d(x, y, z, type = "l", col = "red", lwd = 2, xlab = "X", ylab = "Y", zlab = "Z")

# Add axes and grid
axes3d()
```

```{r fourier transform on a waves}
# Fourier Series for Square Wave approximation
square_wave_fourier <- function(x, n_terms) {
  y <- 0
  for (n in seq(1, n_terms, by = 2)) {
    y <- y + (4 / (pi * n)) * sin(n * x)}
  return(y)}

# create data
x <- seq(0, 6*pi, length.out = 1000)

# Approximate square wave with different numbers of Fourier terms
y_3_terms <- square_wave_fourier(x, 10)   # 3 terms
#y_9_terms <- square_wave_fourier(x, 9)   # 9 terms
#y_99_terms <- square_wave_fourier(x, 500) # 500 terms

# Plot the Fourier Series Approximation
plot(x, y_3_terms, type = "l", col = "blue", lwd = 1, ylim = c(-2, 2), 
     main = "Fourier Series Approximation of Square Wave", xlab = "x", ylab = "y")
#lines(x, y_9_terms, col = "green", lwd = 2)
#lines(x, y_99_terms, col = "red", lwd = 2)

legend("bottomright", legend = c("10  terms"), col = c("blue"), lwd = 1)
```

```{r knot }
t <- seq(0, pi^3, length.out = 1000)

x <- (cos(7 * t)) * cos(2 * t)
y <- (3+cos(7 * t)) * sin(2 * t)
z <- sin(7 * t)

# Plot the 3D cinquefoil knot
plot3d(x, y, z, type = "s", col = "blue", lwd = 2, xlab = "X", ylab = "Y", zlab = "Z")
```

```{r use plot 3d function for other seq}
plottingSeq <- seq(0, 6*pi, length.out = 1500)
# main need is to test different formulas  
formula_str <- "sqrt(plottingSeq)^.05"  # change formula as needed

#HELPER FUNCTION switch text to mathematical formula. A bit recursive 
generateFormula <- function(formula) {
function(x) {eval(parse(text = formula))}}

# HELPER GPT CODE: define formula as a string, create y vals for plot
funkyForm_fn <- generateFormula(formula_str)
formulaValsY <- funkyForm_fn(plottingSeq)

formulaValsZ <- funkyForm_fn(plottingSeq)
# custom colors. Code currently ramps / distributes a color gradient evenly for the n amount of points that exist 
  customPalette <- c("#EDADC7","#A68BA5")
  palette_fn <-colorRampPalette(customPalette) 
  customColors <- palette_fn(length(plottingSeq))
  
  plot3d(formulaValsY,1:1000, 1:1000, type = 'p', col = '#A68BA5',lwd = 1)
```

```{r}
library(ggplot2)

set.seed(42)
df <- data.frame(
  x = runif(100),
  y = runif(100),
  group = sample(1:10, 100, replace = TRUE)
)

ggplot(df, aes(x, y, group = group)) +
  geom_polygon(aes(fill = group), color = "white") +
  theme_void() +
  scale_fill_viridis_d()

```

```{r}
library(aRtsy)

set.seed(123)
canvas_collatz(colors = c("#ff5733", "#33c1ff", "#f5f242"), 100)

```

```{r}
# A trefoil knot
open3d()
theta <- seq(0, 2*pi, length.out = 25)
knot <- cylinder3d(
center = cbind(
sin(theta) + 2*sin(2*theta),
2*sin(3*theta),
cos(theta) - 2*cos(2*theta)),
e1 = cbind(
cos(theta) + 4*cos(2*theta),
6*cos(3*theta),
sin(theta) + 4*sin(2*theta)),
radius = 0.8,
closed = TRUE,
color = "green")
shade3d(addNormals(subdivision3d(knot, depth = 1)))
```
it'd be funny if there was a tag at the end of my code 

           _
       .__(.)< (MEOW)
        \___)   
 
           _
       .__(.)< (**QUACK!**)
        \___)   
 
kermit 

 (+)(+)                     
/      \  
\ -==- / 
 \    / 
<\/\/\/> 
/      \
 Kermit
 
   WW
  /__\
 | oo |      _WWWWW_
(|_()_|)    /  o o  \
  \__/    (|  __O__  |)
 /|\/|\     \ \___/ /
||||||||    /-------\
||||||||   |=========|
||||||||   |=========|
    Bert & Ernie
    
        (o)(o)
       /  ()  \
     _ \ '--' / _ 
    { '-`'"'"`-' }
     `"/      \"`
       \      /
      _/  /\  \_  
    {   /  \   }
      `"`    `"`


          _.-"""-._
         /  _   _  \
        /  (9) (9)  \
       /_,         ,_\
       | \         / |
_      \  \._____./  /  __
\`\     \   \___/   / _|  \
 \ `\   /\         /\ \   /
  |  `\/ /`'-----'`\ \/  /
  |_|\/ /           \   /
  /    /|           |\_/
  \___/ |           | \
   \ .  |           |  \
    \|  |           |  |
     |  `.         .'  |
     \    `-.___.-'    /
     `\       |       /'
       `\     |     /'
    .-.-.`\   |   /'.-.-.
   (,(,(,`^   |   ^`,),),)
    '-'-'-----`-----'-'-'
```{r}
library(ggforce)

set.seed(123)
df <- data.frame(
  x = runif(100),
  y = runif(100),
  group = sample(1:5, 100, replace = TRUE)
)

ggplot(df, aes(x = x, y = y)) +
  geom_voronoi_tile(aes(group = group, fill = group), color = "black") +
  theme_void() +
  scale_fill_manual(limits = NULL)

```

```{r}
library(ggplot2)
library(gganimate)

# Create data with multiple time points
set.seed(42)
n <- 100
df <- data.frame(
  x = runif(n),
  y = runif(n),
  group = sample(1:5, n, replace = TRUE),
  time = rep(1:20, each = n / 20)
)

# Animate the points moving
p <- ggplot(df, aes(x = x, y = y, color = as.factor(group))) +
  geom_point(size = 3) +
  scale_color_viridis_d() +
  theme_void() +
  transition_time(time) + 
  ease_aes('linear')

# Save animation
animate(p, nframes = 100, width = 800, height = 800)
anim_save("generative_animation.gif")

```
Use the rgl() library to plot 3 dimensional pupil data over time. 
  How to do this 
- 5 different colors for the different SNR's 
- use the dilation (AU) as the radius of the circle over time 
 x = time
 y = AU
 z = radius?  
```{r}
library(rgl)
t <- seq(-1, 1, length.out = 1000)

x <- 1:12  # trial length 
y <-  # let this be dilation in AU  
z <- t 

# Plot the 3D cinquefoil knot
plot3d(x,z,y, type = "s", col = "blue", lwd = 2, xlab = "time", ylab = "A.U.", zlab = "radius")
```


